---
title: 'Bios 6301: Assignment 2'
output: pdf_document
author: Yi Zuo
---

*(informally) Due Tuesday, 18 September, 1:00 PM*

50 points total.

This assignment won't be submitted until we've covered Rmarkdown.
Create R chunks for each question and insert your R code appropriately.
Check your output by using the `Knit PDF` button in RStudio.

1. **Working with data** In the `datasets` folder on the course GitHub repo, you will find a file called `cancer.csv`, which is a dataset in comma-separated values (csv) format. This is a large cancer incidence dataset that summarizes the incidence of different cancers for various subgroups. (18 points)

    1. Load the data set into R and make it a data frame called `cancer.df`. (2 points)
```{r}
setwd("/Volumes/GoogleDrive/My Drive/Vanderbilt/1st Semester 2018-2019/BIOS6301 IntroStatComp/Homework_YZ/Assignment 2")

cancer.df<-read.csv("cancer.csv")
head(cancer.df)

```
    2. Determine the number of rows and columns in the data frame. (2)
```{r}
dim(cancer.df)
# 42120 rows and 8 columns
```
    3. Extract the names of the columns in `cancer.df`. (2)
```{r}
colnames(cancer.df)
```
    4. Report the value of the 3000th row in column 6. (2)
```{r}
cancer.df[3000,6]
```
    5. Report the contents of the 172nd row. (2)
```{r}
cancer.df[172,]
```
    6. Create a new column that is the incidence *rate* (per 100,000) for each row.(3)
```{r}
cancer.df<-within(cancer.df,{rate=incidence/population*100000})
```
    7. How many subgroups (rows) have a zero incidence rate? (2)
```{r}
nrow(cancer.df[cancer.df$rate==0,])
```
    8. Find the subgroup with the highest incidence rate.(3)
```{r}
cancer.df[which(cancer.df$rate==max(cancer.df$rate)),]
```

2. **Data types** (10 points)

    1. Create the following vector: `x <- c("5","12","7")`. Which of the following commands will produce an error message? For each command, Either explain why they should be errors, or explain the non-erroneous result. (4 points)

            max(x): the result would be "7", since x is a character vector, and "7" is the highest first digit among "5", and "1".
            sort(x): the result would be "12" "5" "7", since x is a character vector, and `sort` would give the ascending order of the vector.
            sum(x): this would give an error message, since x is a character vector, and can not be summed.

    2. For the next two commands, either explain their results, or why they should produce errors. (3 points)

            y <- c("5",7,12): the result is "5" "7" "12", because numeric value is more flexible than character value, and when they are input together, numeric value is transformed to character value.
            y[2] + y[3]: this would give you an error, because you can not add the elements in the character vector. 

    3. For the next two commands, either explain their results, or why they should produce errors. (3 points)

            z <- data.frame(z1="5",z2=7,z3=12): this would generate a data frame with three variables of `z1`, `z2`, and `z3`, and their values are `"5"`, `7`, and `12`, respectively.
            z[1,2] + z[1,3]: this would calculate the sum of the elements in the first row second column, and first row third column, which is 19.

3. **Data structures** Give R expressions that return the following matrices and vectors (*i.e.* do not construct them manually). (3 points each, 12 total)

    1. $(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)$
```{r}
f1<-function(n){
  c(seq(1,n,1),seq(n-1,1,-1))
}
f1(8)
```
    2. $(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)$
```{r}
f2<-function(n){
  output<-NULL
  for (i in 1:n){
    output<-c(output,rep(i,i))
  }
  return(output)
}
f2(5)
```
    3. $\begin{pmatrix}
      0 & 1 & 1  \\
      1 & 0 & 1  \\
      1 & 1 & 0  \\
    \end{pmatrix}$
```{r}
f3<-function(n){
  matrix<-matrix(rep(1,n^2),n,n)
  for(i in 1:n){
    matrix[i,i]<-0
  }
  return(matrix)
}
f3(3)
```
    4. $\begin{pmatrix}
      1 & 2 & 3 & 4 \\
      1 & 4 & 9 & 16 \\
      1 & 8 & 27 & 64  \\
      1 & 16 & 81 & 256 \\
      1 & 32 & 243 & 1024  \\
    \end{pmatrix}$
```{r}
f4<-function(n){
  df<-matrix(rep(0,4*n),n,4)
  for (i in 1:n){
    df[i,]<-c(1^i,2^i,3^i,4^i)  
  }
 return(df)
}
f4(5)
```
4. **Basic programming** (10 points)

    1. Let $h(x,n)=1+x+x^2+\ldots+x^n = \sum_{i=0}^n x^i$. Write an R program to calculate $h(x,n)$ using a `for` loop. (5 points)

```{r}
q4.1<-function(x,n){
  h<-0
  for (i in 0:n){
    h<-h+x^i
  }
  return(h)
}
```
    2. If we list all the natural numbers below 10 that are multiples of 3 or 5, we get 3, 5, 6 and 9. The sum of these multiples is 23. Write an R program to perform the following calculations. (5 points)

        1. Find the sum of all the multiples of 3 or 5 below 1,000. (3, [euler1])
```{r}
sum<-0
k<-1
while(T){
  if (k%%3 == 0 | k%%5 == 0) sum<-sum+k
  k<-k+1
  if(k==1000) break
}
sum
```
        1. Find the sum of all the multiples of 4 or 7 below 1,000,000. (2)
```{r}
sum<-0
k<-1
while(T){
  if (k%%4 == 0 | k%%7 == 0) sum<-sum+k
  k<-k+1
  if(k==1000000) break
}
sum
```
    3. Each new term in the Fibonacci sequence is generated by adding the previous two terms. By starting with 1 and 2, the first 10 terms will be $(1, 2, 3, 5, 8, 13, 21, 34, 55, 89)$. Write an R program to calculate the sum of the first 15 even-valued terms. (5 bonus points, [euler2])
```{r}
q4.3<-function(m){
  sum<-2
  n1<-1
  n2<-2
  k<-1
  while(T){
  n<-n1+n2
  if (n%%2==0){
    sum<-sum+n
    k<-k+1
  }
  n1<-n2
  n2<-n
  if (k==m) break}
  return(sum)
}
q4.3(15)
```
Some problems taken or inspired by projecteuler.

[euler1]:https://projecteuler.net/problem=1
[euler2]: https://projecteuler.net/problem=2
