---
title: 'Bios 6301: Assignment 6'
author: "Yi Zuo"
output:
  html_document:
    df_print: paged
---

28/30 Good job Yi! - cole h.

*Due Thursday, 15 November, 1:00 PM*

$5^{n=day}$ points taken off for each day late.

30 points total.

Submit a single knitr file (named `homework8.rmd`), along with a valid PDF output file. Inside the file, clearly indicate which parts of your responses go with which problems (you may use the original homework document as a template). Add your name as `author` to the file's metadata section. Raw R code/output or word processor files are not acceptable.

Failure to name file `homework8.rmd` or include author name may result in 5 points taken off.

### Question 1 ###

**15 points**

Install the `readxl` package and run the following

```{r}
fn <- 'icd10.xlsx'
if(file.access(fn, mode = 4) == -1) {
    url <- "https://www.cdc.gov/nhsn/xls/icd10-pcs-pcm-nhsn-opc.xlsx"
    download.file(url, destfile = fn)
}
dat <- readxl::read_excel(fn, sheet = 2)
```

1. Show the class of `dat`. (1 point)

```{r}
class(dat)
```

2. Show the methods available for objects of the given class (if there are multiple classes, show methods for all classes). (3 points)

```{r}
methods(class = class(dat)[1])
methods(class = class(dat)[2])
methods(class = class(dat)[3])
```

3. If you call `print(dat)`, what print method is being dispatched? (1 point)

```{r}
# print(dat)
# print.default(dat)
```

<span style="color:orange"> The default print method is dispatched. </span>

(-1) It's actually not necessarily the default print method. Consider that  the first method available for the object is `tbl_df`, and there exists a `print.tbl_df` method that will be dispatched.

4. Set the class of `dat` to be a data.frame. (1 point)

```{r}
class(dat) <- "data.frame"
```

5. If you call `print(dat)` again, what print method is being dispatched? (1 point)

```{r}
# print(dat)
```

<span style="color:orange"> The `data.frame` print method is dispatched. </span>

Define a new generic function `nUnique` with the code below.

```{r}
nUnique <- function(x) {
    UseMethod('nUnique')
}
```

6. Write a default method for `nUnique` to count the number of unique values in an element. (2 points)

```{r}
nUnique.default <- function(x){
  return(length(unique(x)))
}

```

7. Check your function (2 points)

```{r}
nUnique(letters) # should return 26
nUnique(sample(10, 100, replace = TRUE)) # should return 10 (probably)
```

8. Write a data.frame method for `nUnique` to operate on data.frame objects.
This version should return counts for each column in a data.frame. (2 points)

```{r}
nUnique.data.frame <- function(x){
  z <- sapply(1:ncol(x), function(i) length(unique(x[,i])))
  return(z)
}

```


9. Check your function (2 points)

```{r}
nUnique(dat)
```

### Question 2 ###

**15 points**

Programming with classes.  The following function will generate random patient information.

```{r}
makePatient <- function() {
  vowel <- grep("[aeiou]", letters)
  cons <- grep("[^aeiou]", letters)
  name <- paste(sample(LETTERS[cons], 1), sample(letters[vowel], 1), sample(letters[cons], 1), sep='')
  gender <- factor(sample(0:1, 1), levels=0:1, labels=c('female','male'))
  dob <- as.Date(sample(7500, 1), origin="1970-01-01")
  n <- sample(6, 1)
  doa <- as.Date(sample(1500, n), origin="2010-01-01")
  pulse <- round(rnorm(n, 80, 10))
  temp <- round(rnorm(n, 98.4, 0.3), 2)
  fluid <- round(runif(n), 2)
  list(name, gender, dob, doa, pulse, temp, fluid)
}
```

1. Create an S3 class `medicalRecord` for objects that are a list with the named elements `name`, `gender`, `date_of_birth`, `date_of_admission`, `pulse`, `temperature`, `fluid_intake`. Note that an individual patient may have multiple measurements for some measurements.  Set the RNG seed to `8` and create a medical record by taking the output of `makePatient`.  Print the medical record, and print the class of the medical record. (5 points)

```{r}

mrf <- function(){
  set.seed(8)
  mr <- makePatient()
  names(mr)<-c("name","gender","date_of_birth", "date_of_admission", "pulse", "temperature", "fluid_intake")
  class(mr) <- "medicalRecord"
  return(mr)
}

mr <- mrf()
mr
```


2. Write a `medicalRecord` method for the generic function `mean`, which returns averages for pulse, temperature and fluids. Also write a `medicalRecord` method for `print`, which employs some nice formatting, perhaps arranging measurements by date, and `plot`, that generates a composite plot of measurements over time.  Call each function for the medical record created in part 1. (5 points)

(-1) You probably should plotting all three values of temperature, fluid intake, and pulse on the same graph unless the units are all related in some unusual way, or there's some relationship you are interested in. These values all are measured in quite distinct units, and in a visualization and interpretation sense, should not be present on the same graph (see for instance that you cannot observe much variability in fluid intake since it's on an improper scale, but would on its own scale).

```{r}

# mean function
mean.medicalRecord <- function(x){
  cat(sprintf("Average for pulse: %f\nAverage for temperature: %f\nAverage for fluids: %f", mean(x$pulse), mean(x$temperature),mean(x$fluid_intake)), "\n")
}

mean(mr)

# print function
print.medicalRecord <- function(x) {
 
  date_admission <- x$date_of_admission[order(x$date_of_admission)]
  pulse <- x$pulse[order(x$date_of_admission)]
  temperature <- x$temperature[order(x$date_of_admission)]
  fluid <- x$fluid_intake[order(x$date_of_admission)]
  cat(sprintf("name: %s\ngender: %d\ndate of birth: %s\ndate of admission: %s\npulse: %d\ntemperature: %f\nfluid intake: %f\n", x$name, x$gender, x$date_of_birth, date_admission, pulse, temperature, fluid), "\n")
}

mr

# plot function

plot.medicalRecord <- function(x){
  
  pulse <- x$pulse
  temperature <- x$temperature
  fluid <- x$fluid_intake
  Value <- c(pulse, temperature, fluid)
  Index <- 1:length(pulse)
  Lab <- c(rep("pulse",length(pulse)),rep("temperature",length(temperature)),rep("fluid",length(fluid)))
  
  df <- data.frame(Index,Value,Lab)
  
  library(ggplot2)
  
  ggplot(data=df,aes(x=Index,y=Value,group=Lab)) + geom_line(aes(color=Lab)) 
  
}

plot(mr)


```

3. Create a further class for a cohort (group) of patients, and write methods for `mean` and `print` which, when applied to a cohort, apply mean or print to each patient contained in the cohort. Hint: think of this as a "container" for patients.  Reset the RNG seed to 8 and create a cohort of ten patients, then show the output for `mean` and `print`. (5 points)

```{r}

set.seed(8)
patients <- replicate(10,list(makePatient()))

class(patients) <- "Patients"

# mean
mean.Patients <- function(x){
  pulse <- sapply(1:10, function(i) mean(x[[i]][[5]]))
  temperature <- sapply(1:10, function(i) mean(x[[i]][[6]]))
  fluid <- sapply(1:10, function(i) mean(x[[i]][[7]]))

  message_pulse <- sprintf("Mean pulse for the subject %d: %f",1:10,pulse)
  message_temp <- sprintf("Mean temperature for the subject %d: %f",1:10,temperature)
  message_fluid <- sprintf("Mean fluid for the subject %d: %f",1:10,fluid)
  return(c(message_pulse,message_temp, message_fluid))
}

mean(patients)

# print
print.Patients <- function(x){
  
  # get the number of visits from each person
  n_visit <- sapply(1:10, function(i) length(x[[i]][[4]]))
  
  # store the admission date info
  date_admission <- sapply(1:10, function(i) x[[i]][[4]][order(x[[i]][[4]])])
  
  # pulse
  pulse <- sapply(1:10, function(i) x[[i]][[5]][order(x[[i]][[4]])])
  temperature <- sapply(1:10, function(i) x[[i]][[6]][order(x[[i]][[4]])])
  fluid <- sapply(1:10, function(i) x[[i]][[7]][order(x[[i]][[4]])])
  
  # print
  sapply(1:10, function(i) cat(sprintf("name: %s\ngender: %d\ndate of birth: %s\ndate of admission: %s\npulse: %d\ntemperature: %f\nfluid intake: %f\n", x[[i]][[1]], x[[i]][[2]], x[[i]][[3]], date_admission[[i]], pulse[[i]], temperature[[i]], fluid[[i]]), "\n"))
  
  
}

patients


```
